---
title: "Neural Networks"
output: html_notebook
---

This illustrates neural networks. 
Ref: https://www.r-bloggers.com/fitting-a-neural-network-in-r-neuralnet-package/
Read the file and check for NAs.

```{r}
kuiper <- read.csv("Kuiper.dat")
summary(kuiper)
apply(kuiper,2,function(x) sum(is.na(x)))
```

Now, check the data for outliers
```{r}
plot(kuiper$Price,kuiper$Mileage) # note the outliers
```

Now, exlude the outliers, and select columns for ANN
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.
```{r}
ks1 <- select(kuiper,Price,Mileage,Liter:Leather)
ks2 <- filter(ks1,Price < 52000)
summary(ks2)
```

Let us now divide the set into two. Really, we should divide it into three parts, but for now, we will use a Training and Test set.

```{r}
index <- sample(1:nrow(ks2),round(0.8*nrow(ks2)))
train <- ks2[index,]
test <- ks2[-index,]
summary(train)
summary(test)
```

Data standardization is an important part of ANN. In this example, we will do it as for the most part the data is well constructed.

Let us now run a simple regression model. The results will serve as a base case to test against ANN.
```{r}
lm.fit <- lm(Price~.,data=train)
summary(lm.fit)
pr.lm <- predict(lm.fit,test)
summary(pr.lm)
anova(pr.lm)
mse.lm <- sum((pr.lm-test$Price)^2/nrow(test))
mse.lm
```

Let us now call the neural network on the training set
```{r}
nn <- nnet(Price~.,data=train,size=3)
plot(nn)
```

Now, install the NeuralNetTools packages to plot the network
```{r}
#library(NeuralNetTools)
plotnet(nn,pos.col='darkgreen',neg.col='darkblue',alpha.val=0.7,rel.rsc=15,
circle.cex=10,cex=1.4,
    circle.col='brown')
plotnet(nn,wts.only=T)
```
To determine relative importance of input variables
```{r}
garson(nn)
```

We could also use the neuralnet package
```{r}
library(neuralnet)
nn2 <- neuralnet(Price~Mileage+Liter+Doors+Cruise+Sound+Leather,data=train,hidden=5,rep=1)
plotnet(nn2)
```
