---
title: "Missing Data"
output: html_notebook
---
Here is the link to the presentation: https://docs.google.com/a/kent.edu/presentation/d/10cY3vZghJ4KVybpAO5kWw-e-3bG7PD4D--ZV6KzfUt0/edit?usp=sharing.

Here is a paper that you might find interesting: https://www.jstatsoft.org/article/view/v045i03



Install the requisite packages
```{r}
install.packages("VIM")
install.packages("mice")
```

Read the example file "Equity"
```{r}
equity <- read.csv("Equity.dat",na.strings = ".")
summary(equity)
```

Explore the missing values in this dataset
```{r}
equity[complete.cases(equity),]
equity[!complete.cases(equity),]
sum(is.na(equity$DEBTINC))
mean(is.na(equity$DEBTINC))
```

Now, lets use finer tools to identify patterns of missing data
Use the tabular output from the mice package
```{r}
library("mice")
md.pattern(equity)
```

Use the visual output from the VIM package
```{r}
library("VIM")
aggr(equity,prop=F,numbers=T)
matrixplot(equity)
marginplot(equity[c("CLAGE","MORTDUE")], pch=c(20),
           col=c("darkgray", "red", "blue"))
```

Use correlations
```{r}
x <- as.data.frame(abs(is.na(equity))) # code missing values as 1
head(equity,n=5)
head(x,n=5)

# now extract variables with missing values
y <- x[which(apply(x,2,sum)>0)] # find sum over columns
cor(y)

# create a dataset with only complete cases
z <- na.omit(equity)

```

Now, let's try imputation
```{r}
library(mice)
imp <- mice(equity,seed=1234)
fit <- with(imp,lm(BAD~DEBTINC))
pooled <- pool(fit)
summary(pooled)
```

Look at the imputed object
```{r}
imp
imp$imp$MORTDUE  #look at each object's imputed value
data1 <- complete(imp,action=1) # use 1-5 for determing the imputed set
data1
```

Compare this with complete case analysis
```{r}
fitcc <- lm(BAD~DEBTINC, data=z)
summary(fitcc)
```

Now, let us do Informative Missing
Create two new variables, and run the model on these new variables
```{r}
library(dplyr)
equity <- mutate(equity,DEBTINCorMean = ifelse(is.na(DEBTINC),33.7799,DEBTINC),
       Ind= ifelse(is.na(DEBTINC),1,0))
fit2 <- lm(BAD~DEBTINCorMean+Ind,equity)
fit1 <- lm(BAD~DEBTINC,equity)
summary(fit2)
summary(fit1)
```
