---
title: "Non-linear Models"
output: html_notebook
---

This notebook illustrates the following techniques:
Principal component analysis
Cluster Analysis
Decision / Classification trees
```{r}
toypc <- read.csv("Toy Princomp.dat")
summary(toypc)
```
```{r}
cor(toypc)
```

```{r}
fit <- princomp(toypc[,-3],cor=T)
summary(fit)
```
```{r}
loadings(fit) # pc loadings
plot(fit,type="lines") # scree plot
fit$scores # the principal components
biplot(fit)
```
Now, run a regression of Y on x,z, and Y on Comp.1, Comp.2
```{r}
fit1 <- lm(y~x+z,data=toypc)
fit2 <- lm(toypc$y~fit$scores)
summary(fit1)
summary(fit2)
```

Now, open up the Olymp88sas.dat dataset
```{r}
olymp <- read.csv("Olymp88sas.dat")
summary(olymp)
```
Let's use the "psych" package to run PCA here.
```{r}
library("psych")
fa.parallel(toypc[,-3],fa="pc")
```

We could also use the principal function to do this
```{r}
fit3 <- principal(toypc[,-3],nfactors = 2,scores=T)
fit3
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
